---
title: "Linear Regression Project"
author: "Alana Willis"
date: "6/19/2020"
output: html_document
---


Initial EDA to see trends between variables. Win Percentage, Free Throw Percentage, Average Points Per Game are possible response variables. 

```{r}
library("tidyverse")
nba <- read.csv("http://www.stat.cmu.edu/cmsac/sure/materials/data/regression_projects/nba_team_season_summary.csv")

# Created games per season and average points per game. 
nba <- nba %>%
  mutate(games = wins + losses, avg_points = total_points/games, 
         avg_turnovers= turnovers/games, 
         avg_fouls= personal_fouls/games, 
         avg_assists = assists/games, 
         avg_rebounds = total_rebounds/games, 
         asd = score_differential/games, 
         avg_steals= steals/games, 
         avg_blocks= blocks/games)
table(nba$games)

#Histograms to check for normal distributions.
nba %>%
  ggplot(aes(x = win_perc)) +
  geom_histogram()

nba %>%
  ggplot(aes(x = free_throw_attemps)) +
  geom_histogram()

nba %>%
  ggplot(aes(x = free_throw_perc)) +
  geom_histogram()

nba %>%
  ggplot(aes(x = avg_points)) +
  geom_histogram()

nba %>%
  ggplot(aes(x = asd)) +
  geom_histogram()

nba %>%
  ggplot(aes(x = avg_turnovers)) +
  geom_histogram()


#Scatter plots to check relationship between two variables. 
nba %>%
ggplot(aes(y = win_perc)) +
  geom_point(aes(x = field_goal_perc), stat = "identity")
nba %>%
ggplot(aes(y = win_perc)) +
  geom_point(aes(x = free_throw_perc), stat = "identity")
nba %>%
ggplot(aes(y = win_perc)) +
  geom_point(aes(x = three_point_perc), stat = "identity")

nba %>%
  ggplot(aes(y = free_throw_perc)) +
  geom_point(aes(x = personal_fouls), stat = "identity")
nba %>%
  ggplot(aes(y = free_throw_attemps)) +
  geom_point(aes(x = personal_fouls), stat = "identity")

nba %>%
  ggplot(aes(y = win_perc)) +
  geom_point(aes(x = avg_turnovers), stat = "identity")
nba %>%
  ggplot(aes(y = avg_points)) +
  geom_point(aes(x = avg_turnovers), stat = "identity")
```


Correlation Matrix. 
```{r}
# Win Percentage 
library(ggcorrplot)
nba_corr <- nba %>%
  dplyr::select(win_perc, free_throw_perc, 
                three_point_perc, field_goal_perc, 
                two_point_perc, avg_points, avg_turnovers, avg_fouls, score_differential)
nba_corr_matrix <- cor(nba_corr)
ggcorrplot(nba_corr_matrix)

round_cor_matrix <- round(cor(nba_corr), 2)
ggcorrplot(round_cor_matrix, 
           hc.order = TRUE,
           type = "lower",
           lab = TRUE)


# Score_Diff and Total Points
nba_corr2 <- nba %>%
  dplyr::select(avg_steals, avg_blocks, avg_rebounds, avg_assists, avg_fouls, avg_points, asd, avg_turnovers)
nba_corr_matrix2 <- cor(nba_corr2)
ggcorrplot(nba_corr_matrix2)

round_cor_matrix2 <- round(cor(nba_corr2), 2)
ggcorrplot(round_cor_matrix2, 
           hc.order = TRUE,
           type = "lower",
           lab = TRUE)


```
Pairs Plots
```{r}
library(GGally)
ggpairs(nba_corr2,
        columns =
          c("asd", "avg_steals",
            "avg_blocks", "avg_rebounds",
            "avg_assists","avg_fouls", "avg_points"))
```
We like avg score differential (asd) because it breaks games into three categories: winners, losers, and close games.
We like avg points per game, it was correlated with many other variables and is an easy to understand variable (interpretations)
We averaged all the variables to account for seasons with less games and to improve interpretability.


Initial Linear Regressions. 
```{r}
nba_lm_1 <- lm(asd ~ avg_rebounds + avg_blocks + avg_steals + avg_fouls  + avg_assists + avg_turnovers, data = nba)
summary(nba_lm_1)
plot(nba_lm_1)

nba_lm_2 <- lm(avg_points ~ )
```
EDA for ASD. 
```{r}
nba %>%
  ggplot(aes(y = avg_points))+
  geom_point(aes(x = asd), stat = "identity")

summary(nba$asd)

nba %>%
  mutate(
    asd_category = case_when(
      asd <= -9 ~ 1,
      asd <= -3 ~ 2,
      asd <= 3 ~ 3,
      asd <= 9 ~ 4, 
      asd > 9 ~ 5,
      TRUE ~ asd
    ))
#1=big loss, 2=loss, 3=close game, 4=win, 5=big win
```


